Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-12 paper/main.tex
Squashed some TODOs and FIXMEs
Started plotting bytes hashed by tahoe

Most of the cloud applications today are highly dependant on the use of
cryptographic hash functions. The distributed file system Tahoe-LAFS is no
exception, and uses hashing algorithms extensively for verification of the
integrity of its files.

In light of the ongoing NIST competition for selecting the new, standard
cryptographic hash function named SHA-3, we tried out the 14 second
round candidates in the Tahoe-LAFS file system using automated tests to measure
the performance differences. Only the immutable file support was tested.

Since the Tahoe-LAFS is programmed in Python, we had to make bindings to the
SHA-3 implementations written in C, before making the necessary changes to the
Tahoe-LAFS code. Several tasks where automated using scripts, to avoid doing
manual work of distributing changes to the test grid.

In general, four of the candidates perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.
By utilizing the performance improvements given by SHA-3, Tahoe-LAFS could
save $5\%$ on upload operations by changing hashing function to the best
performing candidate.

The benchmarking tool SUPERCOP measures performance and drives the development
forwards for the SHA-3 candidates, but we argue that since the SUPERCOP API does
not conform with the NIST SHA-3 API, vital functionality are left out, making
integration and testing with existing software hard.

@@

Lastly, the Open Source community behind Tahoe-LAFS deserves credit for the
immense amount of hours they put on an awesome project, and Thomas Pornin for
providing a working copy of his library of implementations, which saved us a lot
of work.
\vspace{13pt}

Best regards,

\begin{center}
Eirik Haver and Pål Ruud
\vspace{13pt}

December 16, 2010
\end{center}

@@

@@

The sum of the amount of time it took for the underlying hash library to
calculate the hash operations are exhibited in Figures \ref{fig:graph:10001b},
\ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
\ref{fig:graph:11gb}. We call this \emph{time spent hashing}.

The numbers for uploading and downloading the 1GB file are summarized in
Tables \ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}.
BLAKE, \ac{BMW}, Shabal and Simd outperforms \ac{SHA}-256 when both uploading
and downloading a 1GB file. The difference between the slowest (JH) and the
fastest (BLAKE) algorithm is 229.92s when uploading the one gigabyte file.

The third column in Tables \ref{tbl:hashingtimes:put1gb} and
\ref{tbl:hashingtimes:get1gb} contains the time spent doing other operations
than hashing, i.e. $Total time - Time spent hashing$. This number varies
between 227.37s and 252.66s during upload and 205.94s and 225.03s during
download. \emph{Fraction} displays the part of total time the calls to the
hashing libraries take.

@@

\section{\ac{SUPERCOP} results}

@@

\section{Validity of \ac{SUPERCOP} results}

@@

\section{Exclusion of mutable file testing}

@@

@@

\subsection{The use of Cython}

@@

\subsection{Faster implementations}
We do not claim to have tested the fastest possible nor existing version
of most of the candidates. We are certain that someone of more skill and
knowledge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
Python could make significantly faster implementation.

@@

We had hoped to contribute more to the Open Source project Tahoe-LAFS, when
starting on our assignment. It was early evident that implementing a new
capability was out of scope, since creating Python bindings and trying out
new implementations was prioritized.

Four of the candidates has been shown to perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd. The time it takes for
Tahoe-LAFS to upload a one gigabyte file, could improve by $5\%$ if the best
performing SHA-3 candidate, BLAKE, were to be used.

>>>>

@@

Lastly, the Open Source community behind Tahoe-LAFS deserves credit for the
immense amount of hours they put on an awesome project, and Thomas Pornin for
providing a working copy of his library of implementations, which saved us a lot
of work.

@@

\section{Measurements of Tahoe data}
% TODO: Fiks denne tabellen så den ikke overflower
% TODO: Sleng inn samme for download
Table \ref{tbl:tahoe:datauploaded} and describes the distribution of the
different input sizes of data that Tahoe-\ac{LAFS} sends to the hash function
when uploading and downloading the test vectors. E.g. to upload the 100 1-byte
files, 15700 hashes had to be calculated, and about 64\% of these were of size
106 byte.

For every upload Tahoe-\ac{LAFS} does, it hashes a size close to the original
file size twice. This does not happen on downloading the file.

\input{tables/tahoe_byteshashed}

@@

The sum of the times it took for the underlying hash library to calculate the
hash operations are exhibited in Figures \ref{fig:graph:10001b},
\ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
\ref{fig:graph:11gb}. We call this \emph{time spent hashing}.

The numbers for uploading and downloading of one 1GB file are summarized in
Tables \ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}.

@@

\section{\ac{SUPERCOP} results}
% Time gain of removing ``double'' SHA-256. Double only in digest().

@@

\section{Validity of \ac{SUPERCOP} results}
% Mention that some SUPERCOP candidates are missing UPDATE-functionality.

@@

\section{Exclusion of mutable file testing}
% Implications of only testing immutable files? (mutable files also
%   include directories)
% Adding a capability to support new hash functions in tahoe. (Our
%   current way breaks backwards compatibility)

@@

\subsection{Small files}
For small files, 1B and 1KB, we can clearly see that both \ac{SHA}-256 and
double \ac{SHA}-256 greatly outperforms the \ac{SHA}-3 candidate
implementations, which is unexpected if the time Tahoe-\ac{LAFS} do mainly
realy on the filesize.

\subsection{Large inputs}

\subsection{The candidates}

\subsection{Optimizing results}

@@

\subsection{The use of Cython}
% What if we had used another coding scheme? - does Cython lower the
%   candidates performance?

@@

\subsection{Faster implementations}
We do not claim to have tested the fastest possible nor existing version
of most of the candidates. We are certain that someone of more skill and
knolwedge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
Python could make significantly faster implementation.

@@

We had hoped to contribute more to the Open Source project Tahoe-LAFS, when
starting on our assignment. It was early evident that implementing a new
capability was out of scope, since creating Python bindings for and trying out
new implementations was prioritized.
